A few stickleback suffice for the transport of alleles to new lakes

### 2019-11-17 review-responses.tex

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

We are writing to submit a revised version of our manuscript, 
``A few stickleback suffice for the transport of alleles to new lakes''.
Many thanks to the reviewers for thoughtful and helpful comments,
to which we have responded in detail.
We agree that additional simulations may well be informative and interesting,
but the amount of work that these would entail is beyond our capability at the moment,
and we think that the additional information added would probably be marginal.
We appreciate being explicitly given the option to omit them.
Otherwise, we think that the revisions have helped to clarify and generally improve the manuscript.
We hope you agree.

The responses to reviews includes page and line numbers that refer to the resubmitted manuscript;
we are also uploading a color-coded pdf that shows differences from the previously submitted version.

### 2019-11-18 second-review-responses.tex

%%%%%%
%%
%%  Don't reorder the reviewer points; that'll mess up the automatic referencing!
%%
%%%%%

\begin{minipage}[b]{2.5in}
  Resubmission Cover Letter \\
    {\it G3}
\end{minipage}
\hfill
\begin{minipage}[b]{2.5in}
    Jared Galloway, \\
    William Cresko, \\
    \emph{and} Peter Ralph \\
  \today
\end{minipage}

\vskip 2em

\noindent
{\bf To the Editor(s) -- }

\vskip 1em

Thanks very much for the prompt and positive decision.
We have added some additional discussion about the two reviewer points
(that previously we had only responded to in the Response to Reviewers),
as requested by the editor.
This was a good idea, as other readers will likely have the same thoughts.

We hope that you agree with our assessment.

\vspace{5em}

\noindent \hspace{4em}
\begin{minipage}{3in}
\noindent
{\bf Sincerely,}

\vskip 2em

{\bf 
Jared Galloway, Bill Cresko, and Peter Ralph.
}\\
\end{minipage}

\vskip 4em

\pagebreak
\setcounter{page}{1}

%%%%%%%%%%%%%%
\reviewersection{AE}

\begin{quote}
I would ask that you add some text to the discussion addressing reviewer points 1.1 and 3.2.
\end{quote}

Here is Reviewer 1 point 1:
\begin{quote}
    \textit{``As expected, the genetic basis of the freshwater phenotype seems to simplify
    as migration increases -- higher rates of migration allow adaptive alleles of
    larger effect to travel more efficiently through the population, even though
    they are deleterious in the ocean.''}
    Very cool. It gives a very intuitive guideline as to which genetic
    architecture of local adaptation to expect given a certain migration rate.
    There is one neat aspect of this result that is perhaps worth better
    emphasizing: detectable $F_{ST}$ peaks at truly adaptive
    alleles can only be very few. If we had replicate runs we could say how many
    true $F_{ST}$ peaks, on average, we see at $M=1$ and $M=10$. Many published studies
    (particularly in marine animals) report dozens if not hundreds of $F_{ST}$ outliers
    throughout the genome; your simulation here suggests that most of them have to
    be false positives of some kind, especially since migration in the sea is
    likely to be high.
\end{quote}

% our previous \reply{
%     We agree with the reviewer that this is an important point,
%     and that our results are consistant with the above hypothesis. 
%     However, to quantify and assert that previous empirical studies 
%     are reporting false-positives would require many 
%     more simulations and analysis in many more situations,
%     which we believe is outside the realm of this paper.
% }

We have added, at \llname{disco11}:

\textit{
We find that, unsurprisingly, sharing of alleles across lakes leads to
a very simple genetic basis -- only tens of loci, at $M \ge 1$.
This suggests that all else being equal, systems with higher migration
should have fewer loci responsible for local adaptation.
However, ``all else is equal'' is rarely plausible:
the actual number of locally adaptive alleles in a particular system
depends on many other factors, including
the strength of selection, degree of pleiotropy, and the genetic basis of the trait.
A different genetic basis could well lead to a much larger number of alleles, even at high migration rates:
but, could these be reliably identified?
Next, we look at how well adaptive alleles are identified with genome scans.
}

% \textit{
% Because the detectable traits at high dispersal were very few,
% empirical studies in the situation of high dispersal which report
% hundreds or thousands of causal loci may benefit from exploring
% false-positive rates in similar modeling studies.
% }

\bigskip

Here is Reviewer point 3.2:
\begin{quote}
If possible, it would be interesting to check what type of mutations - in terms of effect size and in terms of dominance - fall under the 'captured', 'marine' and 'de novo' categories. I could well imagine, for example, that mutations that can stay in the marine populations relatively long are more likely to be recessive. Is this true?
\end{quote}

% our previous \reply{
%     We thought about this also, and did this comparison in some earlier simulations,
%     but any differential representation was not obvious
%     -- there was no large effect --
%     and that we'd need many more simulations to detect anything.
%     So: this seems plausible, but speculatively,
%     it doesn't seem to be a major driver of genetic architecture,
%     but rather a minor adjustment.
% }

We have added, at \llname{disco32}:

\textit{
We explored the possibility of dominance being a good predictor of
how long an adaptive allele can be maintained in the marine environment,
but found no effect strong enough to be seen with our relatively small sample size of adaptive alleles.
}

### 2019-11-18 Stickleback_Paper.tex
revew to responses again

%%%%%%%%%%%
% for responses to reviewers
% set this to show line numbers and include responses to reviews or not
\newif\ifreviewresponses
\reviewresponsestrue  % include them
% \reviewresponsesfalse  % don't include them
\input{review-response-commands}
\newcommand{\responsefile}{review-responses.tex}  % name of the review reponses file

@@

Local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants, 
as can be seen by the mean trait values of \autoref{fig:MeanPhenotype}.
Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat. 
Establishment of adaptive alleles in the old lakes during the first few thousand generations manifests as jumps in the mean trait value -- which move the trait by an amount of order 1 every few hundred generations (\autoref{fig:phenotype_time}). \revpoint{3}{4}
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats. 
As expected, increasing migration rate decreased differentiation between habitats: as seen in \autoref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.  \revpoint{1}{1}

@@

To investigate in more depth how locally adaptive alleles found in the original lakes are shared among newly derived lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation. 
To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes. 
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat. 
(So, ``captured'' alleles as defined above are pre-existing, but not all pre-existing alleles are captured.)
\autoref{fig:MPFAI}A shows the distribution of the number of these alleles across generations. At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones. 
At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other, but now the new lakes adapt using pre-existing alleles present in the original set of lakes. 
Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$. As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (\autoref{fig:MPFAI}C). Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.

\autoref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%. At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
\autoref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes. 
As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean. \revpoint{1}{1}

@@

@@

What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake per generation 
are the only rates of dispersal at which a large percentage of peaks (above $F_{ST} = 0.05$) 
actually lie on top of regions which impact trait differences.
Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{ST}$ peaks. 
In other words, with high rates of dispersal between population, $F_{ST}$ peaks may reliably identify causal loci, 
but not all causal loci will appear as $F_{ST}$ peaks.
It's important to note that $F_{ST}$ in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} was calculated after pooling together all lakes and all marine individuals.
For comparison, Figures~\ref{fig:fst_5lakes} \& \ref{fig:Power_FP_5lakes} shows $F_{ST}$ calculated with individuals only in the first fifth of the range (5 lakes and corresponding marine habitat).
These show a substantially lower true positive rate, as population-specific noise swamps any signal between the lakes.

>>>>

%%%%%%%%%%%
% for responses to reviewers
% set this to show line numbers and include responses to reviews or not
\newif\ifreviewresponses
\reviewresponsestrue  % include them
% \reviewresponsesfalse  % don't include them
\newcommand{\responsefile}{second-review-responses.tex}  % name of the review reponses file
\input{review-response-commands}

@@

Local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants, 
as can be seen by the mean trait values of \autoref{fig:MeanPhenotype}.
Figures~\ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat. 
Establishment of adaptive alleles in the old lakes during the first few thousand generations manifests as jumps in the mean trait value -- which move the trait by an amount of order 1 every few hundred generations (\autoref{fig:phenotype_time}). \revpoint{3}{4}
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites (about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats. 
As expected, increasing migration rate decreased differentiation between habitats: as seen in \autoref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.

@@

To investigate in more depth how locally adaptive alleles found in the original lakes are shared among newly derived lakes, as well as how they spread to the new lakes, we defined and tracked the distribution of ``pre-existing freshwater adapted alleles'' at the beginning of each generation. 
To be considered in this category, an allele must participate in the genetic basis of local adaptation for at least one of the original lakes. 
Concretely, these are any non-neutral mutations whose frequency is above 50\% in at least one original lake and below 50\% in the marine habitat. 
(So, ``captured'' alleles as defined above are pre-existing, but not all pre-existing alleles are captured.)
\autoref{fig:MPFAI}A shows the distribution of the number of these alleles across generations. At $M = 0.01$, each lake has a private set of about 10 mutations nearly fixed in that lake but not elsewhere: new lakes independently acquire new adaptive alleles rather than pre-existing ones. 
At $M = 0.1$, we again observe the original lakes adapting nearly independently from each other, but now the new lakes adapt using pre-existing alleles present in the original set of lakes. 
Concurrently, the average marine individual carries $\approx 2$ pre-existing freshwater adapted alleles, standing variation which was nearly absent at $M=0.01$. As migration rate increases past this, the total number of pre-existing freshwater adapted alleles declines (\autoref{fig:MPFAI}C). Interestingly, the frequency of these alleles in the ocean
stays relatively constant across the reasonable rates of gene flow.
We explored the possibility of dominance being a good predictor of
how long an adaptive allele can be maintained in the marine environment,
but found no effect strong enough to be seen with our relatively small sample size of adaptive alleles.
\llabel{disco32}

\autoref{fig:MPFAI}B shows the distribution, through time, of the mean \emph{percentage} of currently-defined freshwater adapted alleles that each genome in each of the populations carries. 
If all individuals across lakes carried the same set of alleles determining their trait value, this would be 100\%. At the lowest migration rate ($M = 0.01$), each genome in the original lakes have almost exactly~$1/25^{th}$ of the total number of pre-existing freshwater adapted alleles -- this is because each one of the 25 lakes has adapted with a unique set of alleles. 
Since these are \emph{pre-existing} alleles, the value is zero for introduced lakes. 
\autoref{fig:MPFAI}A shows us that at $0.1$ migrants per lake per generation and above, the average individual across the new lakes has nearly the same amount of pre-existing freshwater adapted alleles as individuals across the old lakes. 
As expected, the genetic basis of the freshwater phenotype seems to simplify as migration increases -- higher rates of migration allow adaptive alleles of larger effect to travel more efficiently through the population, even though they are deleterious in the ocean. \revpoint{1}{1}
% Because the detectable traits at high dispersal were very few, empirical studies in the situation of high dispersal which report hundreds or thousands of causal loci may benefit from exploring false-positive rates in similar modeling studies. \llabel{disco11}

@@

We find that, unsurprisingly, sharing of alleles across lakes leads to
a very simple genetic basis -- only tens of loci, at $M \ge 1$.
This suggests that all else being equal, systems with higher migration
should have fewer loci responsible for local adaptation.
However, ``all else is equal'' is rarely plausible:
the actual number of locally adaptive alleles in a particular system
depends on many other factors, including
the strength of selection, degree of pleiotropy, and the genetic basis of the trait. \llabel{disco11}
A different genetic basis could well lead to a much larger number of alleles, even at high migration rates:
but, could these be reliably identified?
Next, we look at how well adaptive alleles are identified with genome scans.

@@

What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake per generation 
are the only rates of dispersal at which a large percentage of peaks (above $F_{ST} = 0.05$) 
actually lie on top of regions which impact trait differences.
Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{ST}$ peaks. 
In other words, with high rates of dispersal between population, $F_{ST}$ peaks may reliably identify causal loci, 
but not all causal loci will appear as $F_{ST}$ peaks.
It's important to note that $F_{ST}$ in Figures~\ref{fig:Fst} and \ref{fig:Power_FP} was calculated after pooling together all lakes and all marine individuals.
For comparison, Figures~ \ref{fig:Power_FP_5lakes} and~\ref{fig:fst_5lakes} show $F_{ST}$ calculated with individuals only in the first fifth of the range (5 lakes and corresponding marine habitat).
These show a substantially lower true positive rate, as population-specific noise swamps any signal between the lakes.
